Anxiety: a decision-theoretic perspective

### 2019-01-13 manuscript/manuscript.tex
sunday writing (pt 2)

1. Open with definition of avoidance. Follow \citep{Arnaudova2017} in definitions, i.e.
the basic idea that avoidance is the distancing of oneself from real or perceived
threat.
2. Describe cliff walking environment. Need to discuss that we are expanding
on the result originally described in \citep{Gaskett2003}.
3. With increasing pessimism, agent further distances self from cliff. Note that
this is the optimal policy under those set of beliefs about the world.
4. This can explain how avoidance is maintained. Even with luck and chance in the
the world, these sets of beliefs actually lead to and sustain these behavioral
policies.

In discussion of section, need to also mention other relevant findings. Specifically,
need to describe the approach-avoidance findings of Bach (2014, 2017) and the
complex task findings of Sheynin et al., 2015.

@@

@@

In the third section of the results, we discuss approach-avoidance conflict.
This covers tasks such as the balloon analog risk task (BART) and the flight
initiation distance task, which has been found to correlate with anxiety. The
hallmark of these sorts of tasks is the tradeoff between increased reward vs.
the possibility of an aversive outcome (e.g. balloon pop, predation). Discuss
previous findings, i.e. earlier leave times.

We can show that this is easily explained by our model. A lack of belief of
future control predicts exiting while things are still good. (We see this in
Figure 3). The optimal policy in these environments is an earlier leaving time.

This can also be used to explain behavioral inhibition \citep{bach2015, khemka2017}.
The idea is slightly different than what is advocated. The claim here is that
in anxious individuals, the Q-values should be closer together, and thus there
should be longer waiting time. This requires demonstrating that decisions are
not made in advance; only at the time of decision making. This is something
I think we will have to argue. (Need to double check the literature here.)

\begin{figure}
  \centerline{%
    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/03_fid.png}}%
  }
  \caption{\textbf{Flight Initiation Distance}}
  \par remember to add parameters
\end{figure}

>>>>

A second symptom central to all anxiety disorders is avoidance \citep{dsm5,
Krypotos2015, Arnaudova2017}, or actions taken to increase or maintain the temporal,
physical and/or mental distance between an individual and a threatening situation.
Although avoidance is a natural response to danger, excessive avoidance as is
observed in anxiety disorders is can be highly disruptive to everyday functioning
\cite{Salter2004} and may work to sustain anxious appraisals and behavior. For
example, an individual with social anxiety may avoid attending large social events
for fear of public embarrassment only to never learn that such worry is not justified.

In the laboratory, avoidance has been measured in a variety of ways. In animal
models of anxiety, anxious rodents are found to exhibit avoidance learning more
quickly and maintain avoidance for longer \citep{servatius2008}. In studies of
human avoidance, anxiety has been found to correlate with the maintenance of
increased distance from threat, such as virtual predators \citep{Bach2014, Bach2017,
Sheynin2014}.

% Arnaudova2017 makes reference to some spider approach studies that might be
% interesting to try and track down. Essential idea is that even though
% participants are totally safe, may not touch the glass. The task is the
% behavioral avoidance task. (e.g. Siegel & Weinberger, 2012, emotion).
% the nice part about this is that participants are objectively safe.

There is a long history of models attempting to explain avoidance learning. The
fundamental challenge is explaining how an action can be reinforced in the
absence of an outcome. Two-factor theory was the explanation for a long while,
but suffered from a number of noteworthy criticisms \citep{Krypotos2015}.
Recent work has revitalized two-factor theory and integrated it with reinforcement
learning \citep{Moutoussis2008, Maia2010}, and in the process addressing several
criticisms of two-factor theory. As has been discussed elsewhere \citep{Moutoussis2017},
such accounts only explain the maintenance of avoidance in short timescales, but
do not explain why resistant to change and why they do not go away in general.

% Need to go into more detail than this, need to actually make this a coherent
% explanation, but this can wait for now.

@@

In our second result, we demonstrate how a breakdown in control can lead to
sustained policies of avoidance. We do so in the cliff-walking environment
\citep{SuttonBarto1998, SuttonBarto2018, Gaskett2003}. Here an agent learns to navigate from
an initial starting position to an end state. The shortest distance is along a
cliff edge. If the agent steps off the cliff, it incurs a large penalty. Under
complete control (objective or perceived), agent in this environment is under no
threat despite the presence of the cliff's edge. As in the first simulation, this
is simply because, in a fully controllable world, an agent need not ever interact
with the aversive state. In fact, an agent can decide to walk straight up to the
aversive state and, insofar it does not choose to take one step further, will
never be in any actual danger.

The long-run state values under different beliefs are plotted in Figure 2. As
expected, under optimistic beliefs (Figure 1b) the entire environment sans negative
state takes on a positive value (close to the positive state itself). As this belief
breaks down, we observe a different pattern in the long-run state values. When
future occupancy of the negative state becomes a non-zero probability, the undesired
action transition exerts influence on action-outcome calculations, back-propagating,
and thereby infecting antecedent states. As the agent's belief becomes increasingly
pessimistic, the more the worst possible action exerts influence and the more
negative, on average, the state-space grows. Most importantly, because these are
the long-run state values, the policy is the optimal policy; in other words, this
is the long-run policy an agent will take.

@@

Next, we turn our attention towards a related phenomenon: a bias towards avoidance
in approach-avoidance conflict. In environments with correlated risk and rewards,
a bias towards avoidance can ultimately result in fewer expected rewards and worse
long-run returns. Returning to the example above, an individual with social anxiety
may forgo social situations to avoid the risk of social embarrassment at the expense
of friendships, etc. This is another way in which pathological anxiety can be
disruptive to everyday functioning.

Experimentally, there is a long history of measuring behavior during
approach-avoidance conflict in individuals with anxiety (citation, maybe yin-yang anxiety paper).
One popular paradigm is the balloon analog risk task (BART) \citep{Lejuez2002},
which has been shown to correlate with anxiety \cite{Maner2007, Giorgetta2012}.
In the task, a participant pumps a balloon with air. For each pump, the participant
earns points. If she pumps too much, the balloon pops and she loses all points.
The participant must decide within each episode when to stop pumping. Leave time
is the measure of behavior, with earlier leave times in anxiety suggesting a
greater tendency towards avoidance.

\begin{figure}
  \centerline{%
    \resizebox{1.0\textwidth}{!}{\includegraphics[trim={0 0 0 0},clip]{../figures/03_fid.png}}%
  }
  \caption{\textbf{BART}}
  \par remember to add parameters
\end{figure}

Our sequential model of decision making easily explains these results. A schematic
of the task is presented in Figure 3. Under optimistic beliefs, points should
travel back from the mean. Under pessimistic assumptions, however, different
degrees of the worst outcome backpropagate. At a certain point, the perceived cost
of staying outweighs the guaranteed gains of leaving. Thus, patients with anxiety
leave earlier.

We can explain similar findings with the flight initiation distance \citep{Mobbs2018,
Mobbs2019}. One thing our model cannot explain is the neural difference between
fast and slow escape. Indeed, our model cannot capture all but perhaps is relevant
for certain calculations. One could imagine these beliefs interacting in such a
way to change the distance calculations that are necessary for mediating between
passive and active defense responses.

This can also be used to explain behavioral inhibition \citep{bach2015, khemka2017}.
The idea is slightly different than what is advocated. The claim here is that
in anxious individuals, the Q-values should be closer together, and thus there
should be longer waiting time.
